### 2017-04-04 paper.tex
Add first packet layout draft

\begin{itemize}
\item Latitude and Longitudes can be both positive (North, East) or Negative
(South, West). At the equator, \({0.0001}^\circ\) \footnotemark of
longitude maps to roughly 11 metres, and that distance decreases when moving
towards the poles, which was deemed precise enough for the project's purpose.
To avoid conflicts in floating-point mathematics conventions, and allow packet
encoders to be written for embedded platform without hardware floating point
support, latitude and longitude were designed to be multiplied by \(10,000\)
and transmitted as 32-bit precision signed integers.

\item Altitude is positive for the whole mission duration, and cannot exceed
\(40,000m\). Since a sub-metre precision is not required, and would hardly be
achievable with civilian GNSS receivers, it is encoded as a 16-bit precision
unsigned integer.
\end{itemize}

\footnotetext{
Earth circumference at the equator: \(40\time10^{6}m\).
\(D_{0.0001^{\circ}} = 0.0001\frac{40\time10^{6}}{360}=11.1...m\)
}

>>>>

\begin{itemize}
\item Latitude and Longitudes can be both positive (North, East) or Negative
(South, West). At the equator, \({0.0001}^\circ\) of longitude maps to roughly
11 metres\footnotemark , and that distance decreases when moving towards the
poles, which was deemed precise enough for the project's purpose. To avoid
conflicts in floating-point mathematics conventions, and allow packet encoders
to be written for embedded platform without hardware floating point support,
latitude and longitude were designed to be multiplied by 10,000 and transmitted
as 32-bit precision signed integers.

\item Altitude is above zero for the whole mission duration, and cannot exceed
50,000m. Since a sub-metre precision is not required, and would hardly be
achievable with civilian GNSS receivers, it is encoded as a 16-bit precision
unsigned integer.
\end{itemize}

\footnotetext{
Earth circumference at the equator: \(40\time10^{6}\)m.
\(D_{0.0001^{\circ}} = 0.0001\frac{40\time10^{6}}{360}=11.1...m\)
}

The decision was made during the design process to align packet boundaries with
frame boundaries. This means that every time a new packet is sent through the
radio channel starts at the start of a new frame, instead of being inserted
in the middle of a frame, where the last packet stopped.

This approach leads to a loss in bandwidth: if a packet's length is not a 
multiple of the amount of data that a frame carry, the packet's last frame
will be padded with zeros instead of containing the start of the next packet.
However, it was deemed an acceptable tradeoff because of the issues that
non-aligned (contiguous) packets would present:

\begin{itemize}
\item If a packet's end does not match a frame's end, that frame cannot be
sent until the start of the next packet is known. This meant that the
transmission of most packet's last frames could be delayed for sometimes several
seconds until another packet is processed.

\item In the case where a frame is lost due to channel noise or interferences,
finding the start of the next packet would have been complicated, or required
a synchronisation mechanism similar to the one designed for AHABus frames.
Aligning packets instead makes such a system unnecessary, reducing the packet
header size and the complexity of packet decoders.
\end{itemize}

% TODO: Add diagram explaining aligment
